---
title: "2023-02-13-ExtraFigures"
author: "Hubert Rehrauer / Jose F Carreño"
output: 
  html_document:
    self_contained: true
    includes:
     in_header: !expr system.file("templates/fgcz_header.html", package="ezRun", lib.loc=.libPaths())
    css: !expr system.file("templates/fgcz.css", package="ezRun", lib.loc=.libPaths())
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---


#  {.tabset}

## Overview

```{r, warning=FALSE, include=FALSE}
### Load necessary libraries

library(Seurat)
library(cowplot)
library(ggplot2)
library(dplyr) 
library(ezRun)
```

This report is intended to provide with some extra features for the final manuscript. The analysis reproduced from previously performed pipelines. 

### Set important paths

```{r}
dataDirs <- list(
   day60_CD4 = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_CD4protocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_CD4protocol_SCReport",
     labeledObject = NULL
   ), # Not used for now
###############################################################
   day60_ctrl = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_controlprotocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_controlprotocol_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60-o30306/day60_ctrl"
   ), # This is the second run of the ctrl experiment
   day60_otic = list(
     cellRanger =
"/srv/gstore/projects/p29781/o29804_CellRangerCount_2022-11-04--11-20-44/day60oticdiff_Library_418886_1",
     seuratReport =
"/srv/gstore/projects/p29781/o29804_ScSeurat_2022-11-08--10-01-51/day60oticdiff_Library_418886_1_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60"
   ), # This is the dataset contaminated with neurons
   day60_merged = list(
     cellRanger = NULL,
     seuratReport = NULL,
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2023-01-27-Extended-MergeThenSelect"
   ), # This is the merged dataset
###############################################################
   day30 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o29467_CellRangerCount_2022-09-16--09-26-33/day30oticvesicle_Library_412321",
     seuratReport =
"/srv/gstore/projects/p27889/o29467_ScSeurat_2022-09-22--15-04-32/day30oticvesicle_Library_412321_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-10-04-relabel-day30oticvesicle"
   ), # Not used for now
   day08 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o28146_CellRangerCount_2022-07-11--09-59-01/placodeday8test1",
     seuratReport =
"/srv/gstore/projects/p27889/o28146_ScSeurat_2022-05-19--11-14-04/placodeday8test1_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-08-19-Seurat-relabeled"
   ) # Not used for now
)


resultDir <- "/home/jcarreno/sta426_project/results/2023-02-13-ExtraFigures"
```

### Genes of interest

This list of genes has been obtained from "list for clustering 7.12.2022.xlsx"

```{r}
reference_genes <- list(
  ghc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF',
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5',
'SLC26A5',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3'
), # General Hair Cells: All markers
hc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF'), # Hair Cells
vhc = c(
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5'), #Vestibular Hair Cells
chc = c(
 'SLC26A5',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3',
'TMPRSS3'),  #Cochlear Hair Cells
oep = c(
'EPCAM',
'CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'PAX2',
'PAX8',
'DLX5',
'GBX2',
'JAG1',
'TBX2',
'COL9A2',
'OTOA',
'MYO5C',
'OTOL1',
'USH1C',
'PCDH9'), # Otic epithelium
goi = c(
'STRC',
'CABP2',
'USH2A',
'TMC2',
'CD164L2',
'SYT14',
'KCNH6',
'PCDH15',
'ISL1',
'OTOG',
'TECTAA',
'GJB2',
'GJB6',
'CLDN14',
'SPARCL1',
'MEIS2',
'SOX9',
'SLC26A4',
'OTOGL'), # Genes of interest (included on 13/02/2023)
sup7 = c(
'CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'GATA3',
'MAFB',
'NEUROD1',
'TNTKR3',
'PRPH',
'NTRK3',
'STMN2',
'DCX',
'TUBB3',
'OTX1',
'ZIC1',
'ZIC2',
'PAX6',
'PAX3',
'OTX2',
'TTr',
'CLIC6',
'HTRC2',
'PLTP',
'CLDN3',
'TFAP2A',
'KRT8',
'TP63',
'KRT19',
'KRT5',
'CXCL14',
'DSP',
'KRT18',
'COL17A1',
'TYR',
'TYRP1',
'ECT',
'EDNRB',
'POSTN',
'PPRRX1',
'TWIST1',
'VIM',
'PDGFRA',
'TWIST2',
'TTN',
'PAX7',
'ACTC1',
'TNNTI',
'TNNII'),
sup5 = c(
  'OC90',
  'SOX2',
  'JAG1',
  'OTOL1',
  'OTOA',
  'USH1C',
  'GBX2',
  'TBX2',
  'ISL1',
  'EYA2',
  'FGF8',
  'GATA3',
  'TUBB3',
  'PRPH',
  'NTRK3',
  'NEUROD4',
  'NGFR',
  'NEUROG1',
  'NEUROD1',
  'GADD45G',
  'HES6',
  'INSM1',
  'EPHAS',
  'TLX3',
  'SALL3',
  'MYT1',
  'SHOX2',
  'NEUROG2',
  'PHOX2B',
  'PHOX2A'
)
)
```

### Define colors

```{r}
seuratPal <- scales::hue_pal()(14)
#scales::show_col(seuratPal)
ctColors <- c(
  "EP"=seuratPal[2],
  "NE"=seuratPal[9],
  "MES"=seuratPal[12],
  "NEURONS"="steelblue",
  "OEP"=seuratPal[4],
  "MUSCLE"="gray40",
  "NE-like"=seuratPal[8],
  "ChP"=seuratPal[11],
  "MEL"="gray60",
  "Chondro"="gray80",
  "HC" = "red")
ezShowCol <- function (colours, colorLabels=names(colours), borders = NULL, cex_label = 1, 
          ncol = NULL) 
{
  n <- length(colours)
  ncol <- ncol %||% ceiling(sqrt(length(colours)))
  nrow <- ceiling(n/ncol)
  colours <- c(colours, rep(NA, nrow * ncol - length(colours)))
  colours <- matrix(colours, ncol = ncol, byrow = FALSE)
  old <- par(pty = "s", mar = c(0, 0, 0, 0))
  on.exit(par(old))
  size <- max(dim(colours))
  plot(c(0, size), c(0, -size), type = "n", xlab = "", ylab = "", 
       axes = FALSE)
  rect(col(colours) - 1, -row(colours) + 1, col(colours), 
       -row(colours), col = colours, border = borders)
  if (!is.null(colorLabels)) {
    hcl <- farver::decode_colour(colours, "rgb", "hcl")
    label_col <- ifelse(hcl[, "l"] > 50, "black", "white")
    text(col(colours)[1:length(colorLabels)] - 0.5, -row(colours)[1:length(colorLabels)] + 0.5, colorLabels, 
         cex = cex_label, col = label_col)
  }
}

ezShowCol(ctColors, colorLabels=paste0(names(ctColors), "\n", ctColors))
```

Load the data from the server and identify the source of each dataset

```{r}
day60_otic <- readRDS(file = paste(dataDirs$day60_otic$labeledObject, "/scData.rds",sep = ""))
day60_ctrl <- readRDS(file = paste(dataDirs$day60_ctrl$labeledObject, "/scData.rds",sep = ""))
day30 <- readRDS(file = paste(dataDirs$day30$labeledObject, "/scData.rds",sep = ""))

day60_otic$experiment <- "day60_otic"
day60_ctrl$experiment <- "day60_ctrl"
```

## Merge the datasets

```{r}
day60_otic <- FindVariableFeatures(day60_otic, nfeatures = 2000)
day60_ctrl <- FindVariableFeatures(day60_ctrl, nfeatures = 2000)
```


```{r}
anchors <- FindIntegrationAnchors(object.list = list(day60_otic, day60_ctrl), dims = 1:20, verbose = TRUE)
```

Integrate the selected cells from both datasets:

```{r}
combined <- IntegrateData(anchorset = anchors, dims = 1:20, verbose = TRUE)
```


From our set of genes stablished in the Overview, check how many of these ARE NOT present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(combined)) %>% length()
})
```

From our set of genes stablished in the Overview, check how many of these ARE present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% intersect(rownames(combined)) %>% length()
})
```


### Reclustering

```{r}
combined <- ScaleData(combined, verbose = FALSE)
combined <- FindVariableFeatures(combined, nfeatures = 2000)
combined <- RunPCA(combined, npcs = 30, verbose = FALSE)
ElbowPlot(combined)
```

```{r}
combined <- RunTSNE(combined, reduction = "pca", dims = 1:5)
combined <- RunUMAP(combined, reduction = "pca", dims = 1:20, verbose = FALSE)
```


```{r}
DimPlot(combined, label = TRUE, cols=ctColors)
```

```{r}
DimPlot(combined, group.by = "Condition")
```

```{r, fig.width=12, fig.height=5}
DimPlot(combined, split.by = "Condition", cols=ctColors)
```

## Select cells

```{r}
DefaultAssay(combined) <- "RNA"
combined <- ScaleData(combined, verbose = FALSE, features=rownames(combined))
```

We compute the actual score: 

```{r}
combined <- AddModuleScore(combined, features=reference_genes, name="module")
```

### Plot scores

```{r, fig.width=20, fig.height=15}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(combined, group.by = "clusterLabels", cols=ctColors)
p2 <- FeaturePlot(combined, featNames)
p3 <- VlnPlot(combined, featNames, group.by = "clusterLabels", ncol = 5, cols=ctColors)

upper_row <- plot_grid(p1, p2, ncol = 2)

plot_grid(upper_row, p3, ncol = 1)
```

```{r}
for (i in 1:length(reference_genes)){
  if(i==2){
    print(VlnPlot(combined, featNames[i], pt.size=1, group.by = "clusterLabels", cols=ctColors) + geom_hline(yintercept=0.30) + ggtitle(names(reference_genes)[i]))
  }
  else if (i==5) {
    print(VlnPlot(combined, featNames[i], pt.size=1, group.by = "clusterLabels", cols=ctColors) + geom_hline(yintercept=0.1) + ggtitle(names(reference_genes)[i]))
  } 
  else {
     print(VlnPlot(combined, featNames[i], pt.size=1, group.by = "clusterLabels", cols=ctColors) + ggtitle(names(reference_genes)[i]))
  }
}
```

Use only the cells above the threshold in module 2 and the otic epithelium cells above threshold in module 5:


```{r}
useCell <- (combined$clusterLabels == "OEP" & combined$module5>0.1) | combined$module2 > 0.30 
combined_sel <- combined[ , useCell]
```

### Reclustering of the selected cells

```{r}
combined_sel <- FindVariableFeatures(combined_sel, nfeatures = 2000)
combined_sel <- RunPCA(combined_sel, npcs = 10, verbose = FALSE)
ElbowPlot(combined_sel)
```

```{r}
combined_sel <- RunUMAP(combined_sel, reduction = "pca", dims = 1:5)
```

```{r}
combined_sel <- RunTSNE(combined_sel, reduction = "pca", dims = 1:5)
combined_sel <- FindNeighbors(combined_sel, reduction = "pca", dims = 1:5, k.param = 7)
```

```{r}
combined_sel <- FindClusters(combined_sel, resolution = 0.3)
```

```{r}
DimPlot(combined_sel, label = TRUE)
```

```{r}
DimPlot(combined_sel, group.by = "Condition")
```

```{r}
DimPlot(combined_sel, group.by = "clusterLabels", cols=ctColors)
```

## Module Score

Compute again the score for each set of genes established in the Overview section:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined_sel, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=10, fig.height=10}
rgSet <- lapply(reference_genes, intersect, rownames(combined_sel))
DoHeatmap(combined_sel, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined_sel))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


Cells in cluster 5 are likely to be HC:

```{r}
Idents(combined_sel) <- combined_sel$seurat_clusters %>% recode("5"="HC")
```


HIGHLIGHT VERSION 1: ALL GREEN VS HC IN RED

```{r}
cells.use <- WhichCells(combined_sel, idents = c('HC'))
DimPlot(combined_sel, reduction = "umap", cells.highlight = cells.use, sizes.highlight = 0.5, cols=seuratPal[4]) + NoLegend()
```


HIGHLIGHT VERSION 2: ALL CELLS THEIR RESPECTIVE COLORS & HC RED

```{r}
Idents(combined_sel) <- combined_sel$clusterLabels
combined_sel <- SetIdent(combined_sel, cells = cells.use, value = 'HC')
```


```{r}
DimPlot(combined_sel, cols=ctColors)
```

HIGHLIGHT VERSION 3: ALL GRAY VS HC IN RED

```{r}
DimPlot(combined_sel, reduction = "umap", cells.highlight = cells.use, sizes.highlight = 0.5) + NoLegend()
```

## Mapping HC to original Merged Dataset

```{r, fig.width=12, fig.height=4}
p1 <- DimPlot(combined, cols=ctColors)
combined_reclustered <- SetIdent(combined, cells = cells.use, value = 'HC')
p2 <- DimPlot(combined_reclustered, cols=ctColors)

plot_grid(p1, p2, ncol=2)
```



## Markers Analysis {.tabset}

### General Hair Cells Markers {.tabset}

#### Heatmap

```{r, fig.height=10, fig.width=8}
DoHeatmap(combined_sel, reference_genes$ghc)
```


#### UMAPs

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Dotplot

```{r}
DotPlot(combined_sel, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


#### Violin Plots


```{r}
for (marker in reference_genes$ghc){
  print(VlnPlot(combined_sel, marker, cols=ctColors))
  #tryCatch({print(VlnPlot(combined_sel, marker, cols=ctColors))},
  #        error=function(e){print("ERROR")},
  #        warning=function(w){print(VlnPlot(combined_sel, marker, cols=ctColors))})
}
```



### Hair Cells Markers {.tabset}

#### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined_sel, reference_genes$hc)
```


#### UMAPs

```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Dotplot

```{r}
DotPlot(combined_sel, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


#### Violin Plots


```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$hc){
  print(VlnPlot(combined_sel, marker, cols=ctColors))
  #tryCatch({print(VlnPlot(combined_sel, marker, cols=ctColors))},
  #         error=function(e){print("ERROR")},
  #         warning=function(w){print(paste(marker, "not present"))})
}
```



### Cochlear Hair Cells Markers {.tabset}

#### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined_sel, reference_genes$chc)
```


#### UMAPs

```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Dotplot

```{r}
DotPlot(combined_sel, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


#### Violin Plots


```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$chc){
  print(VlnPlot(combined_sel, marker, cols=ctColors))
  #tryCatch({print(VlnPlot(combined_sel, marker, cols=ctColors))},
  #         error=function(e){print("ERROR")},
  #         warning=function(w){print(paste(marker, "not present"))})
}
```


### Otic Epithelium Markers {.tabset}

#### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined_sel, reference_genes$oep)
```


#### UMAPs

```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Dotplot

```{r}
DotPlot(combined_sel, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


#### Violin Plots


```{r, fig.width=5, fig.height=4}
for (marker in reference_genes$oep){
  print(VlnPlot(combined_sel, marker, cols=ctColors))

  #tryCatch({print(VlnPlot(combined_sel, marker, cols=ctColors))},
  #         error=function(e){print("ERROR")},
  #         warning=function(w){print(paste(marker, "not present"))})
}
```





## Extra Figures (Violin plots, UMAPS) {.tabset}

### Heatmap

```{r, fig.height=10, fig.width=8}
DoHeatmap(combined_sel, reference_genes$goi)
```


### UMAPs

```{r}
for (marker in reference_genes$goi){
  tryCatch({print(FeaturePlot(combined_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


### Dotplot

```{r}
DotPlot(combined_sel, features=reference_genes$goi) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


### Violin Plots


```{r}
for (marker in reference_genes$goi){
  tryCatch({print(VlnPlot(combined_sel, marker, cols=ctColors) + theme(legend.position = 'none'))},
          error=function(e){print("ERROR")},
          warning=function(w){print(VlnPlot(combined_sel, marker, cols=ctColors)+ theme(legend.position = 'none'))})
}
```



## Extra Figures (Dotplot)

Dotplot with genes for supplementary figure 7:

**THIS FIGURE CONTAINS DATA FROM THE COMPLETE MERGED DATASET:**

```{r,fig.width=12 , fig.height=10}
DotPlot(combined_reclustered, features=reference_genes$sup7) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```

**THIS FIGURE CONTAINS DATA ONLY FROM THE SELECTED CELLS:**

```{r,fig.width=8 , fig.height=10}
DotPlot(combined_sel, features=reference_genes$sup7) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```

## Extra Figure (Supplementary 5): Violin plots

```{r}
day30Colors <- c("OE"=seuratPal[4],
                 "EP"=seuratPal[1],
                 "EP6/K"=seuratPal[2],
                 "ONB"=rgb(10, 32, 160, maxColorValue = 255),
                 "NE"=seuratPal[9],
                 "MES"=seuratPal[12])
```


```{r}
DimPlot(day30, group.by = "cellType", cols=day30Colors)
```

```{r}
for (marker in reference_genes$sup5){
  
  tryCatch({print(VlnPlot(day30, marker, group.by = "cellType", idents = c("OE1, OE2, OE3, OE4", "OE5", "ONB") , cols=day30Colors)+ theme(legend.position = 'none'))},
          error=function(e){print("ERROR")},
          warning=function(w){print(VlnPlot(combined_sel, marker, cols=ctColors))})
}
```




## SessionInfo

```{r, echo=FALSE}
saveRDS(combined, file = file.path(resultDir, "scData_combined.rds"))
saveRDS(combined_sel, file = file.path(resultDir, "scData_combined_selected.rds"))
saveRDS(combined_reclustered, file = file.path(resultDir, "scData_combined_reclustered.rds"))

sessionInfo()
```
