---
title: "Combine day60_otic + day60_ctrl"
author: "Hubert Rehrauer / Jose F Carreño"
output: 
  html_document:
    self_contained: true
    includes:
     in_header: !expr system.file("templates/fgcz_header.html", package="ezRun", lib.loc=.libPaths())
    css: !expr system.file("templates/fgcz.css", package="ezRun", lib.loc=.libPaths())
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

#  {.tabset}

## Overview

```{r, warning=FALSE, include=FALSE}
### Load necessary libraries

library(Seurat)
library(cowplot)
library(ggplot2)
library(dplyr) 
library(ezRun)
```

The interactive explorer ist at <https://fgcz-shiny.uzh.ch/fgcz_iSEE/?data=p29781/additional-analyses/2023-01-21-MergeSamples/sce_h5>

This report is intented to address the different options proposed of merging, filtering and clustering the two experiments EXP45 & EXP46 to obtain a pure cluster of Hair Cells.

Important references are:
<<<<<<< HEAD
  *   Dataset integration algorithm: [https://doi.org/10.1016/j.cell.2019.05.031](https://doi.org/10.1016/j.cell.2019.05.031)
  *   Calculate module score reference: [https://doi.org/10.1126%2Fscience.aad0501](https://doi.org/10.1126%2Fscience.aad0501)
  *   Calculate module score tutorial: [Link](https://www.waltermuskovic.com/2021/04/15/seurat-s-addmodulescore-function/#:~:text=Calculate%20the%20average%20expression%20levels,randomly%20selected%20from%20each%20bin.) (I found this particularly helpful to understand the actual implementation of the algorithm in R)
=======
  -   Dataset integration algorithm: [https://doi.org/10.1016/j.cell.2019.05.031](https://doi.org/10.1016/j.cell.2019.05.031)
  -   Calculate module score reference: [https://doi.org/10.1126%2Fscience.aad0501](https://doi.org/10.1126%2Fscience.aad0501)
  -   Calculate module score tutorial: [Link](https://www.waltermuskovic.com/2021/04/15/seurat-s-addmodulescore-function/#:~:text=Calculate%20the%20average%20expression%20levels,randomly%20selected%20from%20each%20bin.) (I found this particularly helpful to understand the actual implementation of the algorithm in R)
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5

The organization of this report is as follows:

*   Overview: Definition of important variables for the report

*   Analysis 1: Select then merge

    *   Select cells with a given gene expression pattern + OEP cells for each dataset independently
    *   Merge both datasets
    *   Reclustering of selected cells
    *   Subsequent marker analysis
        -   General Hair Cells
        -   Hair Cells
        -   Cochlear Hair Cells
        -   Vestibular Hair Cells
        -   Otic Epithelium

*   Analysis 2: Merge then select

    *   Merge both datasets
    *   Select cells with a given gene expression pattern + OEP cells in the merged dataset
    *   Reclustering of selected cells
    *   Subsequent marker analysis
        -   General Hair Cells
        -   Hair Cells
        -   Cochlear Hair Cells
        -   Vestibular Hair Cells
        -   Otic Epithelium

*   Analysis 3: Merge then select (EPCAM+)

    *   Merge both datasets
    *   Select EPCAM+ cells
    *   Reclustering of selected cells
    *   Subsequent marker analysis
        -   General Hair Cells
        -   Hair Cells
        -   Cochlear Hair Cells
        -   Vestibular Hair Cells
        -   Otic Epithelium

*   Analysis 4: Merge then select (SOX2+)

    *   Merge both datasets
    *   Select SOX2+ cells
    *   Reclustering of selected cells
    *   Subsequent marker analysis
        -   General Hair Cells
        -   Hair Cells
        -   Cochlear Hair Cells
        -   Vestibular Hair Cells
        -   Otic Epithelium

*   Analysis 5: Merge then select (POU4F3+)

    *   Merge both datasets
    *   Select POU4F3+ cells
    *   Reclustering of selected cells
    *   Subsequent marker analysis
        -   General Hair Cells
        -   Hair Cells
        -   Cochlear Hair Cells
        -   Vestibular Hair Cells
        -   Otic Epithelium

### Set important paths

```{r}
dataDirs <- list(
   day60_CD4 = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_CD4protocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_CD4protocol_SCReport",
     labeledObject = NULL
   ), # Not used for now
###############################################################
   day60_ctrl = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_controlprotocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_controlprotocol_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60-o30306/day60_ctrl"
   ), # This is the second run of the ctrl experiment
   day60_otic = list(
     cellRanger =
"/srv/gstore/projects/p29781/o29804_CellRangerCount_2022-11-04--11-20-44/day60oticdiff_Library_418886_1",
     seuratReport =
"/srv/gstore/projects/p29781/o29804_ScSeurat_2022-11-08--10-01-51/day60oticdiff_Library_418886_1_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60"
   ), # This is the dataset contaminated with neurons
###############################################################
   day30 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o29467_CellRangerCount_2022-09-16--09-26-33/day30oticvesicle_Library_412321",
     seuratReport =
"/srv/gstore/projects/p27889/o29467_ScSeurat_2022-09-22--15-04-32/day30oticvesicle_Library_412321_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-10-04-relabel-day30oticvesicle"
   ), # Not used for now
   day08 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o28146_CellRangerCount_2022-07-11--09-59-01/placodeday8test1",
     seuratReport =
"/srv/gstore/projects/p27889/o28146_ScSeurat_2022-05-19--11-14-04/placodeday8test1_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-08-19-Seurat-relabeled"
   ) # Not used for now
)
#dataDirs

resultDir <- "/scratch/jcarreno/sta426_project/results"
```

### Genes of interest

This list of genes has been obtained from "list for clustering 7.12.2022.xlsx"

```{r}
reference_genes <- list(
  ghc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF',
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5',
'SLC26A5',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3'
), # General Hair Cells: All markers
hc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF'), # Hair Cells
vhc = c(
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5'), #Vestibular Hair Cells
chc = c(
 'SLC26A5',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3',
'TMPRSS3'),  #Cochlear Hair Cells
oep = c(
'EPCAM',
'CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'PAX2',
'PAX8',
'DLX5',
'GBX2',
'JAG1',
'TBX2',
'COL9A2',
'OTOA',
'MYO5C',
'OTOL1',
'USH1C',
'PCDH9')
)
```

Load the data from the server and we identify the source of each dataset

```{r}
day60_otic <- readRDS(file = paste(dataDirs$day60_otic$labeledObject, "/scData.rds",sep = ""))
day60_ctrl <- readRDS(file = paste(dataDirs$day60_ctrl$labeledObject, "/scData.rds",sep = ""))

day60_otic$experiment <- "day60_otic"
day60_ctrl$experiment <- "day60_ctrl"
```

## Analysis 1: Select then merge

### Compute Modules Scores

We compute the actual score

```{r}
day60_otic <- AddModuleScore(day60_otic, features=reference_genes, name="module")
day60_ctrl <- AddModuleScore(day60_ctrl, features=reference_genes, name="module")
```

#### Plot scores: day60_otic

```{r, fig.width=20, fig.height=20}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(day60_otic, label=TRUE)
p2 <- FeaturePlot(day60_otic, featNames)
<<<<<<< HEAD
p3 <- VlnPlot(day60_otic, featNames, ncol=5)

grid1 <- plot_grid(p1, p2, ncol = 2)

plot_grid(grid1, p3, ncol = 1)
=======
p3 <- VlnPlot(day60_otic, featNames)

plot_grid(p1, p2, p3, ncol = 2)
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
```

```{r}
for (i in 1:length(reference_genes)){
  print(VlnPlot(day60_otic, featNames[i], pt.size=1) + geom_hline(yintercept=0.3) + ggtitle(names(reference_genes)[i]))
}
```

Use only the cells above the threshold and the otic epithelium cells:

```{r}
useCell <- Idents(day60_otic) == "OEP" | day60_otic$module2 > 0.3 
day60_oticSel <- day60_otic[ , useCell]
```

#### Plot scores: day60_ctrl

```{r, fig.width=18, fig.height=15}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(day60_ctrl, label=TRUE)
p2 <- FeaturePlot(day60_ctrl, featNames)
<<<<<<< HEAD
p3 <- VlnPlot(day60_ctrl, featNames, ncol=5)

grid1 <- plot_grid(p1, p2, ncol = 2)

plot_grid(grid1, p3, ncol = 1)
=======
p3 <- VlnPlot(day60_ctrl, featNames)

plot_grid(p1, p2, p3, ncol = 2)
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
```

```{r}
for (i in 1:length(reference_genes)){
  print(VlnPlot(day60_ctrl, featNames[i], pt.size=1) + geom_hline(yintercept=0.3) + ggtitle(names(reference_genes)[i]))
}
```

Use only the cells above the threshold and the otic epithelium cells:

```{r}
useCell <- Idents(day60_ctrl) == "OEP" | day60_ctrl$module2 > 0.3 
day60_ctrlSel <- day60_ctrl[ , useCell]
```

### Merge datasets


```{r}
day60_oticSel <- FindVariableFeatures(day60_oticSel, nfeatures = 2000)
day60_ctrlSel <- FindVariableFeatures(day60_ctrlSel, nfeatures = 2000)
```

From our set of genes stablished in the Overview, check how many of these are not present in our variables genes for day60_otic experiment:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(day60_oticSel)) %>% length()
})
```

```{r}
anchors <- FindIntegrationAnchors(object.list = list(day60_oticSel, day60_ctrlSel), dims = 1:10)
```

Integrate the selected cells from both datasets:

```{r}
combined <- IntegrateData(anchorset = anchors, dims = 1:10)
```

From our set of genes stablished in the Overview, check how many of these ARE NOT present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(combined)) %>% length()
})
```

From our set of genes stablished in the Overview, check how many of these ARE present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% intersect(rownames(combined)) %>% length()
})
```

#### Reclustering

```{r}
#DefaultAssay(combined) <- "RNA"
combined <- ScaleData(combined, verbose = FALSE)
combined <- FindVariableFeatures(combined, nfeatures = 2000)
combined <- RunPCA(combined, npcs = 10, verbose = FALSE)
ElbowPlot(combined)
```

```{r}
combined <- RunUMAP(combined, reduction = "pca", dims = 1:10)
```

```{r}
combined <- RunTSNE(combined, reduction = "pca", dims = 1:10)
combined <- FindNeighbors(combined, reduction = "pca", dims = 1:10, k.param = 9)
```

```{r}
combined <- FindClusters(combined, resolution = 0.5)
```

```{r}
DimPlot(combined, label = TRUE)
```

```{r}
DimPlot(combined, group.by = "Condition")
```

```{r}
DimPlot(combined, group.by = "clusterLabels")
```


```{r}
Idents(combined) <- combined$seurat_clusters %>% recode("11"="hc1", "12"="hc2")
DefaultAssay(combined) <- "RNA"
combined <- ScaleData(combined, verbose = FALSE, features=rownames(combined))
```

### Signature Analysis {.tabset}

Compute again the score for each set of genes established in the Overview section:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=10, fig.height=10}
rgSet <- lapply(reference_genes, intersect, rownames(combined))
DoHeatmap(combined, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```

Cluster 11 & 12 are likely to be HC cluster

```{r}
cells.use <- WhichCells(combined, idents = c('hc1', 'hc2'))
DimPlot(combined, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```


#### General Hair Cells Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined, reference_genes$ghc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Hair Cells Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined, reference_genes$hc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Cochlear Hair Cells Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined, reference_genes$chc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Otic Epithelium Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined, reference_genes$oep)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


## Analysis 2: Merge then select

### Merge the datasets

```{r}
day60_otic <- FindVariableFeatures(day60_otic, nfeatures = 2000)
day60_ctrl <- FindVariableFeatures(day60_ctrl, nfeatures = 2000)
```


```{r}
anchors <- FindIntegrationAnchors(object.list = list(day60_otic, day60_ctrl), dims = 1:20)
```

Integrate the selected cells from both datasets:

```{r}
combined2 <- IntegrateData(anchorset = anchors, dims = 1:20)
```

From our set of genes stablished in the Overview, check how many of these ARE NOT present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(combined2)) %>% length()
})
```

From our set of genes stablished in the Overview, check how many of these ARE present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% intersect(rownames(combined2)) %>% length()
})
```


### Reclustering

```{r}
combined2 <- ScaleData(combined2, verbose = FALSE)
combined2 <- FindVariableFeatures(combined2, nfeatures = 2000)
combined2 <- RunPCA(combined2, npcs = 30, verbose = FALSE)
ElbowPlot(combined2)
```

```{r}
combined2 <- RunUMAP(combined2, reduction = "pca", dims = 1:20)
```


```{r}
DimPlot(combined2, label = TRUE)
```

```{r}
DimPlot(combined2, group.by = "Condition")
```
```{r, fig.width=12, fig.height=6}
DimPlot(combined2, split.by = "Condition")
```

### Compute Modules Scores

```{r}
DefaultAssay(combined2) <- "RNA"
combined2 <- ScaleData(combined2, verbose = FALSE, features=rownames(combined2))
```

We compute the actual score: 

```{r}
combined2 <- AddModuleScore(combined2, features=reference_genes, name="module")
```

#### Plot scores: combined (Analysis 2)

```{r, fig.width=20, fig.height=20}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(combined2, group.by = "clusterLabels")
p2 <- FeaturePlot(combined2, featNames)
<<<<<<< HEAD
p3 <- VlnPlot(combined2, featNames, group.by = "clusterLabels", ncol = 5)
=======
p3 <- VlnPlot(combined2, featNames, group.by = "clusterLabels")
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5

upper_row <- plot_grid(p1, p2, ncol = 2)

plot_grid(upper_row, p3, ncol = 1)
```

```{r}
for (i in 1:length(reference_genes)){
<<<<<<< HEAD
  if(i==2){
    print(VlnPlot(combined2, featNames[i], pt.size=1, group.by = "clusterLabels") + geom_hline(yintercept=0.30) + ggtitle(names(reference_genes)[i]))
  }
  else if (i==5) {
    print(VlnPlot(combined2, featNames[i], pt.size=1, group.by = "clusterLabels") + geom_hline(yintercept=0.1) + ggtitle(names(reference_genes)[i]))
  } 
  else {
     print(VlnPlot(combined2, featNames[i], pt.size=1, group.by = "clusterLabels") + ggtitle(names(reference_genes)[i]))
  }
}
```

Use only the cells above the threshold in module 2 and the otic epithelium cells above threshold in module 5:


```{r}
useCell <- (combined2$clusterLabels == "OEP" & combined2$module5>0.1) | combined2$module2 > 0.30 
=======
  print(VlnPlot(combined2, featNames[i], pt.size=1, group.by = "clusterLabels") + geom_hline(yintercept=0.30) + ggtitle(names(reference_genes)[i]))
}
```

Use only the cells above the threshold and the otic epithelium cells:


```{r}
useCell <- combined2$clusterLabels == "OEP" | combined2$module2 > 0.30 
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
combined2_sel <- combined2[ , useCell]
```

### Reclustering of the selected cells

```{r}
combined2_sel <- FindVariableFeatures(combined2_sel, nfeatures = 2000)
combined2_sel <- RunPCA(combined2_sel, npcs = 10, verbose = FALSE)
ElbowPlot(combined2_sel)
```

```{r}
<<<<<<< HEAD
combined2_sel <- RunUMAP(combined2_sel, reduction = "pca", dims = 1:5)
```

```{r}
combined2_sel <- RunTSNE(combined2_sel, reduction = "pca", dims = 1:5)
combined2_sel <- FindNeighbors(combined2_sel, reduction = "pca", dims = 1:5, k.param = 7)
```

```{r}
combined2_sel <- FindClusters(combined2_sel, resolution = 0.3)
=======
combined2_sel <- RunUMAP(combined2_sel, reduction = "pca", dims = 1:10)
```

```{r}
combined2_sel <- RunTSNE(combined2_sel, reduction = "pca", dims = 1:10)
combined2_sel <- FindNeighbors(combined2_sel, reduction = "pca", dims = 1:10, k.param = 9)
```

```{r}
combined2_sel <- FindClusters(combined2_sel, resolution = 0.5)
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
```

```{r}
DimPlot(combined2_sel, label = TRUE)
```

```{r}
DimPlot(combined2_sel, group.by = "Condition")
```

```{r}
DimPlot(combined2_sel, group.by = "clusterLabels")
```

### Signature Analysis {.tabset}

Compute again the score for each set of genes established in the Overview section:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined2_sel, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=10, fig.height=10}
rgSet <- lapply(reference_genes, intersect, rownames(combined2_sel))
DoHeatmap(combined2_sel, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined2_sel))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


Cells in cluster 8 are likely to be HC:

```{r}
<<<<<<< HEAD
Idents(combined2_sel) <- combined2_sel$seurat_clusters %>% recode("5"="hc")
=======
Idents(combined2_sel) <- combined2_sel$seurat_clusters %>% recode("8"="hc")
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
cells.use <- WhichCells(combined2_sel, idents = c('hc'))
DimPlot(combined2_sel, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```

<<<<<<< HEAD
#### Markers Discovery

##### All vs all

```{r}
markers <- FindAllMarkers(object=combined2_sel, test.use = "wilcox", only.pos=TRUE, assay="RNA")
## Significant markers
markers <- markers[ ,c("gene","cluster","pct.1", "pct.2", "avg_log2FC","p_val_adj")]
#cm <- cm[cm$p_val_adj < 0.05, ]
markers$cluster <- as.factor(markers$cluster)
markers$diff_pct = abs(markers$pct.1-markers$pct.2)
markers <- markers[order(markers$diff_pct, decreasing = TRUE),] ## why would we round here?? %>% mutate_if(is.numeric, round, digits=3)
#ritexl::write_xlsx(markers, path=file.path(resultDir, "posMarkers.xlsx"))
```

```{r heatmap, fig.width=12, fig.height=12}
top5 <- markers %>% group_by(cluster) 
top5 <- slice_max(top5, n = 5, order_by = avg_log2FC)
genesToPlot <- unique(as.character(top5$gene))
genesToPlot <- intersect(genesToPlot, rownames(combined2_sel))

DoHeatmap(combined2_sel, features=unique(genesToPlot))
DotPlot(combined2_sel, features=genesToPlot) + coord_flip()
```

##### HC vs Rest

```{r}

markers <- FindMarkers(object=combined2_sel, ident.1 = "hc", test.use = "wilcox", assay="RNA")
## Significant markers
markers <- ezFrame(gene=rownames(markers), markers)
markers <- markers[ ,c("gene","pct.1", "pct.2", "avg_log2FC","p_val_adj")]
#cm <- cm[cm$p_val_adj < 0.05, ]
markers$diff_pct = abs(markers$pct.1-markers$pct.2)
markers <- markers[order(markers$diff_pct, decreasing = TRUE),] ## why would we round here?? %>% mutate_if(is.numeric, round, digits=3)
#writexl::write_xlsx(markers, path=file.path(resultDir, "hcMarkers.xlsx"))


```

```{r , fig.width=12, fig.height=12}

genesToPlot <- c(rownames(markers[order(-markers$avg_log2FC), ]) %>% head(30),
                 rownames(markers[order(markers$avg_log2FC), ]) %>% head(30)) %>% unique()


DoHeatmap(combined2_sel, features=unique(genesToPlot))
DotPlot(combined2_sel, features=genesToPlot) + coord_flip()
```
=======
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5

#### General Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=10, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$ghc)
```


##### UMAPs

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$hc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Cochlear Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$chc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Otic Epithelium Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$oep)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

## Analysis 3: Merge then EPCAM+

### Merge the datasets

The merge is the same as in Analysis 2, hence it is skipped. The results would be the same

### Select only EPCAM+ cells

```{r}
FeaturePlot(combined2, reduction = "umap", features = "EPCAM", order= TRUE)
```



```{r}
useCell <-  WhichCells(object = combined2, expression = EPCAM > 1.5, slot = 'scale.data')
combined2_sel <- combined2[ , useCell]
```

Show the selected cells:

```{r}
FeaturePlot(combined2_sel, reduction = "umap", features = "EPCAM", order= TRUE)
```


### Reclustering of the selected cells

```{r}
combined2_sel <- FindVariableFeatures(combined2_sel, nfeatures = 2000)
combined2_sel <- RunPCA(combined2_sel, npcs = 20, verbose = FALSE)
ElbowPlot(combined2_sel)
```

```{r}
combined2_sel <- RunUMAP(combined2_sel, reduction = "pca", dims = 1:20)
```

```{r}
combined2_sel <- RunTSNE(combined2_sel, reduction = "pca", dims = 1:20)
combined2_sel <- FindNeighbors(combined2_sel, reduction = "pca", dims = 1:20, k.param = 5)
```

```{r}
combined2_sel <- FindClusters(combined2_sel, resolution = 0.6)
```

```{r}
DimPlot(combined2_sel, label = TRUE)
```

```{r}
DimPlot(combined2_sel, group.by = "Condition")
```

```{r}
DimPlot(combined2_sel, group.by = "clusterLabels")
```

### Signature Analysis {.tabset}

Compute the module score for each set of genes established in the Overview section and only for EPCAM+ cells:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined2_sel, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=10, fig.height=10}
rgSet <- lapply(reference_genes, intersect, rownames(combined2_sel))
DoHeatmap(combined2_sel, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined2_sel))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


Cells in cluster 13 are likely to be HC:

```{r}
Idents(combined2_sel) <- combined2_sel$seurat_clusters %>% recode("13"="hc")
cells.use <- WhichCells(combined2_sel, idents = c('hc'))
DimPlot(combined2_sel, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```


#### General Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=10, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$ghc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$hc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Cochlear Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$chc)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Otic Epithelium Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$oep)
```


##### UMAPs

<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


<<<<<<< HEAD
```{r, fig.width=5, fig.height=4}
=======
```{r}
>>>>>>> 95a7ec9017eac89511e02685aa10ed249ed6d4d5
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

## Analysis 4: Merge then SOX2+

### Merge the datasets

The merge is the same as in Analysis 2 & 3, hence it is skipped. The results would be the same

### Select only SOX2+ cells

```{r}
FeaturePlot(combined2, reduction = "umap", features = "SOX2", order= TRUE)
```



```{r}
useCell <-  WhichCells(object = combined2, expression = SOX2 > 1.5, slot = 'scale.data')
combined2_sel <- combined2[ , useCell]
```

Show the selected cells:

```{r}
FeaturePlot(combined2_sel, reduction = "umap", features = "SOX2", order= TRUE)
```


### Reclustering of the selected cells

```{r}
combined2_sel <- FindVariableFeatures(combined2_sel, nfeatures = 2000)
combined2_sel <- RunPCA(combined2_sel, npcs = 15, verbose = FALSE)
ElbowPlot(combined2_sel)
```

```{r}
combined2_sel <- RunUMAP(combined2_sel, reduction = "pca", dims = 1:15)
```

```{r}
combined2_sel <- RunTSNE(combined2_sel, reduction = "pca", dims = 1:15)
combined2_sel <- FindNeighbors(combined2_sel, reduction = "pca", dims = 1:15, k.param = 5)
```

```{r}
combined2_sel <- FindClusters(combined2_sel, resolution = 0.6)
```

```{r}
DimPlot(combined2_sel, label = TRUE)
```

```{r}
DimPlot(combined2_sel, group.by = "Condition")
```

```{r}
DimPlot(combined2_sel, group.by = "clusterLabels")
```

### Signature Analysis {.tabset}

Compute the module score for each set of genes established in the Overview section and only for EPCAM+ cells:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined2_sel, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=10, fig.height=10}
rgSet <- lapply(reference_genes, intersect, rownames(combined2_sel))
DoHeatmap(combined2_sel, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined2_sel))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


Cells in cluster 18 & 19 are likely to be HC:

```{r}
Idents(combined2_sel) <- combined2_sel$seurat_clusters %>% recode("18"="hc1", "19"="hc2")
cells.use <- WhichCells(combined2_sel, idents = c('hc1', 'hc2'))
DimPlot(combined2_sel, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```


#### General Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=10, fig.width=15}
DoHeatmap(combined2_sel, reference_genes$ghc)
```


##### UMAPs

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$hc)
```


##### UMAPs

```{r}
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Cochlear Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$chc)
```


##### UMAPs

```{r}
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Otic Epithelium Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$oep)
```


##### UMAPs

```{r}
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

## Analysis 5: Merge then POU4F3+

### Merge the datasets

The merge is the same as in Analysis 2 & 3 & 4, hence it is skipped. The results would be the same

### Select only POU4F3+ cells

```{r}
FeaturePlot(combined2, reduction = "umap", features = "POU4F3", order= TRUE)
```



```{r}
useCell <-  WhichCells(object = combined2, expression = POU4F3 > 1.5, slot = 'scale.data')
combined2_sel <- combined2[ , useCell]
```

Show the selected cells:

```{r}
FeaturePlot(combined2_sel, reduction = "umap", features = "POU4F3", order= TRUE)
```


### Reclustering of the selected cells

```{r}
combined2_sel <- FindVariableFeatures(combined2_sel, nfeatures = 2000)
combined2_sel <- RunPCA(combined2_sel, npcs = 8, verbose = FALSE)
ElbowPlot(combined2_sel)
```

```{r}
combined2_sel <- RunUMAP(combined2_sel, reduction = "pca", dims = 1:8)
```

```{r}
#combined2_sel <- RunTSNE(combined2_sel, reduction = "pca", dims = 1:8)
combined2_sel <- FindNeighbors(combined2_sel, reduction = "pca", dims = 1:8, k.param = 5)
```

```{r}
combined2_sel <- FindClusters(combined2_sel, resolution = 0.5)
```

```{r}
DimPlot(combined2_sel, label = TRUE)
```

```{r}
DimPlot(combined2_sel, group.by = "Condition")
```

```{r}
DimPlot(combined2_sel, group.by = "clusterLabels")
```

### Signature Analysis {.tabset}

Compute the module score for each set of genes established in the Overview section and only for EPCAM+ cells:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined2_sel, featNames, order= TRUE)
```

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r, fig.width=8, fig.height=8}
rgSet <- lapply(reference_genes, intersect, rownames(combined2_sel))
DoHeatmap(combined2_sel, unlist(rgSet))
```


Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined2_sel))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


Cells in cluster 1 are likely to be HC:

```{r}
Idents(combined2_sel) <- combined2_sel$seurat_clusters %>% recode("1"="hc")
cells.use <- WhichCells(combined2_sel, idents = c('hc'))
DimPlot(combined2_sel, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```


#### General Hair Cells Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined2_sel, reference_genes$ghc)
```


##### UMAPs

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Hair Cells Markers {.tabset}

##### Heatmap

```{r}
DoHeatmap(combined2_sel, reference_genes$hc)
```


##### UMAPs

```{r}
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



#### Cochlear Hair Cells Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$chc)
```


##### UMAPs

```{r}
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


#### Otic Epithelium Markers {.tabset}

##### Heatmap

```{r, fig.height=8, fig.width=8}
DoHeatmap(combined2_sel, reference_genes$oep)
```


##### UMAPs

```{r}
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined2_sel, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```


##### Dotplot

```{r}
DotPlot(combined2_sel, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```


##### Violin Plots


```{r}
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined2_sel, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```



## SessionInfo

```{r, echo=FALSE}
sessionInfo()
```