---
title: "EPCAM+analysis"
author: "Jose F Carreno"
format: html
embed-resources: true
editor: source
---


```{r, warning=FALSE}
library(Seurat)
library(ggplot2)
library(ggpubr)
```


## Set important paths

```{r}
dataDirs <- list(
  day60 = list(
    cellRanger = "/srv/gstore4users/p29781/o29804_CellRangerCount_2022-11-04--11-20-44/day60oticdiff_Library_418886_1",
    seuratReport = "/srv/gstore4users/p29781/o29804_ScSeurat_2022-11-08--10-01-51/day60oticdiff_Library_418886_1_SCReport",
    labeledObject = "/srv/gstore4users/p29781/additional-analyses/2022-11-28-relabel-day60"
  ),
  day30 = list(
    cellRanger = "/srv/gstore4users/p27889/o29467_CellRangerCount_2022-09-16--09-26-33/day30oticvesicle_Library_412321",
    seuratReport = "/srv/gstore4users/p27889/o29467_ScSeurat_2022-09-22--15-04-32/day30oticvesicle_Library_412321_SCReport",
    labeledObject = "/srv/gstore4users/p27889/additional-analyses/2022-10-04-relabel-day30oticvesicle"
  ),
  day08 = list(
    cellRanger = "/srv/gstore4users/p27889/o28146_CellRangerCount_2022-07-11--09-59-01/placodeday8test1",
    seuratReport = "/srv/gstore4users/p27889/o28146_ScSeurat_2022-05-19--11-14-04/placodeday8test1_SCReport",
    labeledObject = "/srv/gstore4users/p27889/additional-analyses/2022-08-19-Seurat-relabeled"
  )
)
dataDirs

resultDir <- "/scratch/jcarreno/sta426_project/results"
```

## Load the annotated object

```{r}
anno <- readRDS(file = paste(dataDirs$day60$labeledObject, "/scData.rds",sep = ""))
```

## Original UMAP

```{r}
DimPlot(anno, reduction = "umap", group.by = "ident")
```

```{r}
hc <- c("ATOH1", "ANXA4", "GFI1", "CCER2", "POU4F3", "ATOH1",
"MYO7A",
"MYO6",
"PCP4",
"ESPN",
"TMPRSS3",
"BDNF",
"GNG8",
"POU4F3",
"OTOF",
"FSIP1",
"ZBBX",
"SKOR1",
"DNAH5",
"SCL26A5",
"GATA3",
"LMOD3",
"FGF8",
"TMPRSS3",
"INSM1",
"DNM3"
)
```

## Select EPCAM+ cells


```{r}
anno.epcam <- subset(anno, cells = WhichCells(anno, expression = EPCAM > 1.5))
DimPlot(anno.epcam, reduction = "umap", group.by = "ident")
```

### Recluster EPCAM+ cells with Louvain method

```{r, fig.align='center'}
ElbowPlot(anno.epcam)
```

```{r}
anno.epcam <- FindNeighbors(anno.epcam, dims = 1:19, k.param = 5)
anno.epcam <- FindClusters(anno.epcam, algorithm = 1)
anno.epcam$sub_cluster <- as.character(Idents(anno.epcam))
```

```{r}
DimPlot(anno.epcam, reduction = "umap", label = TRUE)
```
```{r, fig.width=15, fig.height=15}
DoHeatmap(anno.epcam, features = hc)
```


### Number of cells per cluster

```{r}
cellCountperCluster <- data.frame(id = Idents(anno.epcam))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```


## UMAP for set of genes

### Otic epithelium markers

```{r}
oep <- c(
'EPCAM',
'CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'PAX2',
'PAX8',
'DLX5',
'GBX2',
'JAG1',
'TBX2',
'COL9A2',
'OTOA',
'MYO5C',
'OTOL1',
'USH1C',
'PCDH9')
```


```{r}
for (marker in oep){
  tryCatch({ print(FeaturePlot(anno.epcam, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print("WARNING")})
}
```

### General Hair Cells Markers


```{r}
ghc <- c('ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF',
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5',
'SCL26A5',
'GATA3',
'LMOD3',
'FGF8',
'TMPRSS3',
'INSM1',
'DNM3'
)
```

```{r}
for (marker in ghc){
  tryCatch({print(FeaturePlot(anno.epcam, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print("WARNING")})
}
```

### All other clusters

```{r}
oc <- c('TTr',
'CLIC6',
'HTRC2',
'PLTP',
'CLDN3',
'TFAP2A',
'KRT8',
'TP63',
'KRT19',
'KRT5',
'CXCL14',
'DSP',
'KRT18',
'COL17A1',
'TYR',
'TYRP1',
'ECT',
'SOX10',
'EDNRB',
'POSTN',
'PPRRX1',
'TWIST1',
'VIM',
'PDGFRA',
'TWIST2',
'TTN',
'PAX7',
'ACTC1',
'MYLPF',
'TNNTI',
'TNNII')
```


```{r}
for (marker in oc){
  tryCatch({print(FeaturePlot(anno.epcam, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print("WARNING")})
}
```

## Heatmap for genes of interest

```{r}
gi1 <- c('CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'ATOH1',
'ANXA4',
'GFI1',
'CCER2',
'POU4F3',
'GATA3',
'MAFB',
'NEUROD1',
'TNTKR3',
'PRPH',
'NTRK3',
'STMN2',
'DCX',
'TUBB3',
'OTX1',
'ZIC1',
'ZIC2',
'PAX6',
'PAX3',
'OTX2')
```


```{r, fig.width=10, fig.height=7, out.width = "100%", out.height = "100%"}
DoHeatmap(anno.epcam, features = gi1)
```

```{r}
gi2 <- c(
  'TTr',
'CLIC6',
'HTRC2',
'PLTP',
'CLDN3',
'TFAP2A',
'KRT8',
'TP63',
'KRT19',
'KRT5',
'CXCL14',
'DSP',
'KRT18',
'COL17A1',
'TYR',
'TYRP1',
'ECT',
'SOX10',
'EDNRB',
'POSTN',
'PPRRX1',
'TWIST1',
'VIM',
'PDGFRA',
'TWIST2',
'TTN',
'PAX7',
'ACTC1',
'TNNTI',
'TNNII'
)
```


```{r, fig.width=10, fig.height=7, out.width = "100%", out.height = "100%"}
DoHeatmap(anno.epcam, features = gi2)
```

## Expression dot plot 

```{r}
DotPlot(anno.epcam, features=gi1) + coord_flip()
```
```{r}
DotPlot(anno.epcam, features=gi2) + coord_flip()
```


## Show new cluster of HC in the original UMAP


```{r}
cells.use <- WhichCells(anno.epcam, idents = '13')
anno <- SetIdent(anno, cells = cells.use, value = 'Reclustered HC')
```

```{r}
DimPlot(anno, reduction = "umap", group.by = "ident")
```


```{r}
DimPlot(anno, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```

## Violin plots

### All clusters

```{r, fig.width=10, fig.height=7, out.width = "100%", out.height = "100%"}
for (marker in oep){
  print(VlnPlot(anno, marker))
}
```

To check the structure of the EPCAM cluster and confirm the distribution of the violin plot:

```{r}
RidgePlot(anno, features = "EPCAM", ncol = 2)
```



```{r}
for (marker in ghc){
  tryCatch({print(VlnPlot(anno, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print("WARNING")})
}
```


### Only EPCAM+ cells

```{r, fig.width=10, fig.height=7, out.width = "100%", out.height = "100%"}
for (marker in oep){
  print(VlnPlot(anno.epcam, marker))
}
```


```{r, fig.width=10, fig.height=7, out.width = "100%", out.height = "100%"}
for (marker in ghc){
  tryCatch({print(VlnPlot(anno.epcam, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print("WARNING")})
}
```