---
title: "Combine day60_otic + day60_ctrl"
author: "Hubert Rehrauer / Jose F Carreño"
output: 
  html_document:
    self_contained: true
    includes:
      in_header: !expr system.file("templates/fgcz_header.html", package="ezRun", lib.loc=.libPaths())
    css: !expr system.file("templates/fgcz.css", package="ezRun", lib.loc=.libPaths())
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

#  {.tabset}


```{r, warning=FALSE, include=FALSE}
### Load necessary libraries

library(Seurat)
library(cowplot)
library(ggplot2)
library(dplyr) 
library(ezRun)
```


## Overview

We select the cells based on them being OEP or expressing HC markers. As a second steps the two samples are combined. Clustering the integrated data gives two clusters that are hair cells (12 and 13). We name them "hc1" and "hc2".

Important are the tabs "Markers" where you find markers for each cluster, [posMarkers.xlsx](posMarkers.xlsx) and "HC cluster markers" where you find markers differentiating the two HC clusters, [hcMarkers.xlsx](hcMarkers.xlsx).

The interactive explorer ist at <https://fgcz-shiny.uzh.ch/fgcz_iSEE/?data=p29781/additional-analyses/2023-01-21-MergeSamples/sce_h5>

The organization of this report is as follows:

-   Overview: Definition of important variables for the report

-   Analysis 1: Select then merge

    -   Select cells with a given gene expression pattern + OEP cells for each dataset independently
    -   Merge both datasets
    -   Reclustering of selected cells
    -   Subsequent marker analysis

-   Analysis 2: Merge then select

    -   Merge both datasets
    -   Select cells with a given gene expression pattern + OEP cells in the merged dataset
    -   Reclustering of selected cells
    -   Subsequent marker analysis

-   Analysis 3: Merge then select (EPCAM+)

    -   Merge both datasets
    -   Select EPCAM+ cells
    -   Reclustering of selected cells
    -   Subsequent marker analysis

-   Analysis 4: Merge then select (SOX2+)

    -   Merge both datasets
    -   Select SOX2+ cells
    -   Reclustering of selected cells
    -   Subsequent marker analysis

-   Analysis 5: Merge then select (POU4F3+)

    -   Merge both datasets
    -   Select POU4F3+ cells
    -   Reclustering of selected cells
    -   Subsequent marker analysis

### Set important paths

```{r}
dataDirs <- list(
   day60_CD4 = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_CD4protocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_CD4protocol_SCReport",
     labeledObject = NULL
   ), # Not used for now
###############################################################
   day60_ctrl = list(
     cellRanger =
"/srv/gstore/projects/p29781/o30306_CellRangerCount_2022-12-20--11-08-02/IEOday60_controlprotocol",
     seuratReport =
"/srv/gstore/projects/p29781/o30306_ScSeurat_2022-12-20--22-03-42/IEOday60_controlprotocol_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60-o30306/day60_ctrl"
   ), # This is the second run of the ctrl experiment
   day60_otic = list(
     cellRanger =
"/srv/gstore/projects/p29781/o29804_CellRangerCount_2022-11-04--11-20-44/day60oticdiff_Library_418886_1",
     seuratReport =
"/srv/gstore/projects/p29781/o29804_ScSeurat_2022-11-08--10-01-51/day60oticdiff_Library_418886_1_SCReport",
     labeledObject =
"/srv/gstore/projects/p29781/additional-analyses/2022-11-28-relabel-day60"
   ), # This is the dataset contaminated with neurons
###############################################################
   day30 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o29467_CellRangerCount_2022-09-16--09-26-33/day30oticvesicle_Library_412321",
     seuratReport =
"/srv/gstore/projects/p27889/o29467_ScSeurat_2022-09-22--15-04-32/day30oticvesicle_Library_412321_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-10-04-relabel-day30oticvesicle"
   ), # Not used for now
   day08 = list(
     cellRanger =
"/srv/gstore/projects/p27889/o28146_CellRangerCount_2022-07-11--09-59-01/placodeday8test1",
     seuratReport =
"/srv/gstore/projects/p27889/o28146_ScSeurat_2022-05-19--11-14-04/placodeday8test1_SCReport",
     labeledObject =
"/srv/gstore/projects/p27889/additional-analyses/2022-08-19-Seurat-relabeled"
   ) # Not used for now
)
#dataDirs

resultDir <- "/scratch/jcarreno/sta426_project/results"
```

### Genes of interest

This list of genes has been obtained from "list for clustering 7.12.2022.xlsx"

```{r}
reference_genes <- list(
  ghc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF',
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5',
'SCL26A6',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3'
), # General Hair Cells: All markers
hc = c(
'ATOH1',
'MYO7A',
'MYO6',
'PCP4',
'ANXA4',
'GFI1',
'ESPN',
'TMPRSS3',
'BDNF',
'CCER2',
'GNG8',
'POU4F3',
'OTOF'), # Hair Cells
vhc = c(
'FSIP1',
'ZBBX',
'SKOR1',
'DNAH5'), #Vestibular Hair Cells
chc = c(
 'SCL26A6',
'GATA3',
'LMOD3',
'FGF8',
'INSM1',
'DNM3',
'TMPRSS3'),  #Cochlear Hair Cells
oep = c(
'EPCAM',
'CDH1',
'SOX2',
'SIX1',
'OC90',
'SOX10',
'FBXO2',
'LMX1A',
'PAX2',
'PAX8',
'DLX5',
'GBX2',
'JAG1',
'TBX2',
'COL9A2',
'OTOA',
'MYO5C',
'OTOL1',
'USH1C',
'PCDH9')
)
```

## Analysis 1: Select then merge

### Compute Modules Scores

First we load the data from the server and we identify the source of each dataset

```{r}
day60_otic <- readRDS(file = paste(dataDirs$day60_otic$labeledObject, "/scData.rds",sep = ""))
day60_ctrl <- readRDS(file = paste(dataDirs$day60_ctrl$labeledObject, "/scData.rds",sep = ""))

day60_otic$experiment <- "day60_otic"
day60_ctrl$experiment <- "day60_ctrl"
```

We compute the actual score: \[REFERENCE\]

```{r}
day60_otic <- AddModuleScore(day60_otic, features=reference_genes, name="module")
day60_ctrl <- AddModuleScore(day60_ctrl, features=reference_genes, name="module")
```

#### Plot scores: day60_otic

```{r, fig.width=20, fig.height=20}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(day60_otic, label=TRUE)
p2 <- FeaturePlot(day60_otic, featNames)
p3 <- VlnPlot(day60_otic, featNames)

plot_grid(p1, p2, p3, ncol = 2)
```

```{r}
for (i in 1:length(reference_genes)){
  print(VlnPlot(day60_otic, featNames[i], pt.size=1) + geom_hline(yintercept=0.3) + ggtitle(names(reference_genes)[i]))
}
```

Use only the cells above the threshold and the otic epithelium cells:

```{r}
useCell <- Idents(day60_otic) == "OEP" | day60_otic$module2 > 0.3 
day60_oticSel <- day60_otic[ , useCell]
```

#### Plot scores: day60_ctrl

```{r, fig.width=18, fig.height=15}
featNames <- paste0("module", 1:length(reference_genes))
p1 <- DimPlot(day60_ctrl, label=TRUE)
p2 <- FeaturePlot(day60_ctrl, featNames)
p3 <- VlnPlot(day60_ctrl, featNames)

plot_grid(p1, p2, p3, ncol = 2)
```

```{r}
for (i in 1:length(reference_genes)){
  print(VlnPlot(day60_ctrl, featNames[i], pt.size=1) + geom_hline(yintercept=0.3) + ggtitle(names(reference_genes)[i]))
}
```

Use only the cells above the threshold and the otic epithelium cells:

```{r}
useCell <- Idents(day60_ctrl) == "OEP" | day60_ctrl$module2 > 0.3 
day60_ctrlSel <- day60_ctrl[ , useCell]
```

### Merge datasets

```{r}
day60_oticSel <- FindVariableFeatures(day60_oticSel, nfeatures = 2000)
day60_ctrlSel <- FindVariableFeatures(day60_ctrlSel, nfeatures = 2000)
```

From our set of genes stablished in the Overview, check how many of these are not present in our selected cells from day60_otic experiment:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(day60_oticSel)) %>% length()
})
```

```{r}
anchors <- FindIntegrationAnchors(object.list = list(day60_oticSel, day60_ctrlSel), dims = 1:10)
```

Integrate the selected cells from both datasets:

```{r}
combined <- IntegrateData(anchorset = anchors, dims = 1:10)
```

From our set of genes stablished in the Overview, check how many of these ARE NOT present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% setdiff(rownames(combined)) %>% length()
})
```

From our set of genes stablished in the Overview, check how many of these ARE present in the "integrated" assay:

```{r}
sapply(reference_genes, function(rg){
  rg %>% intersect(rownames(combined)) %>% length()
})
```

#### Reclustering

```{r}
DefaultAssay(combined) <- "RNA"
combined <- ScaleData(combined, verbose = FALSE)
combined <- FindVariableFeatures(combined, nfeatures = 3000)
combined <- RunPCA(combined, npcs = 10, verbose = FALSE)
ElbowPlot(combined)
```

```{r}
combined <- RunUMAP(combined, reduction = "pca", dims = 1:10)
```

```{r}
combined <- RunTSNE(combined, reduction = "pca", dims = 1:10)
combined <- FindNeighbors(combined, reduction = "pca", dims = 1:10, k.param = 5)
```

```{r}
combined <- FindClusters(combined, resolution = 0.7)
```

```{r}
DimPlot(combined, label = TRUE)
```

```{r}
DimPlot(combined, group.by = "Condition")
```

```{r}
DimPlot(combined, group.by = "clusterLabels")
```

### Signature Analysis

Compute again the score for each set of genes established in the Overview section:

```{r, fig.width=10, fig.height=10}
featNames <- paste0("module", 1:length(reference_genes))
FeaturePlot(combined, featNames, order= TRUE)
```

#### Heatmaps

Plot the set of genes stablished in the Overview and that are present in the integrated object:

```{r}
rgSet <- lapply(reference_genes, intersect, rownames(combined))
DoHeatmap(combined, unlist(rgSet))
```

Heatmap of the General Hair Cells genes

```{r}
DoHeatmap(combined, reference_genes$ghc)
```

Heatmap of the Hair Cells genes

```{r}
DoHeatmap(combined, reference_genes$hc)
```

Heatmap of the Cochlear Hair Cell genes

```{r}
DoHeatmap(combined, reference_genes$chc)
```

Heatmap of the Vestibular Hair Cell genes

```{r}
DoHeatmap(combined, reference_genes$vhc)
```

Heatmap of the Otic Epithelium genes

```{r}
DoHeatmap(combined, reference_genes$oep)
```

#### UMAPS

UMAPs of the General Hair Cells genes

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

UMAPs of the Hair Cells genes

```{r}
for (marker in reference_genes$hc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

UMAPs of the Cochlear Hair Cells genes

```{r}
for (marker in reference_genes$chc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

UMAPs of the Vestibular Hair Cells genes

```{r}
for (marker in reference_genes$vhc){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

UMAPs of the Otic Epithilium genes

```{r}
for (marker in reference_genes$oep){
  tryCatch({print(FeaturePlot(combined, reduction = "umap", features = marker, order= TRUE))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

#### Dot plots

Dotplot of the General Hair Cells genes

```{r}
DotPlot(combined, features=reference_genes$ghc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```

Dotplot of the Hair Cells genes

```{r}
DotPlot(combined, features=reference_genes$hc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```

Dotplot of the Cochlear Hair Cells genes

```{r}
DotPlot(combined, features=reference_genes$chc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  

```

Dotplot of the Vestibular Hair Cells genes

```{r}
DotPlot(combined, features=reference_genes$vhc) + coord_flip() + theme(axis.text.x = element_text(size = 8))  

```

Dotplot of the Otic Epithilium genes

```{r}
DotPlot(combined, features=reference_genes$oep) + coord_flip() + theme(axis.text.x = element_text(size = 8))  
```

#### Violin plots

Violin plot of the General Hair Cells genes

```{r}
for (marker in reference_genes$ghc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

Violin plot of the Hair Cells genes

```{r}
for (marker in reference_genes$hc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

Violin plot of the Cochlear Hair Cells genes

```{r}
for (marker in reference_genes$chc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

Violin plot of the Vestibular Hair Cells genes

```{r}
for (marker in reference_genes$vhc){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

Violin plot of the Otic Epithilum genes

```{r}
for (marker in reference_genes$oep){
  tryCatch({print(VlnPlot(combined, marker))},
           error=function(e){print("ERROR")},
           warning=function(w){print(paste(marker, "not present"))})
}
```

Number of cells per cluster:

```{r}
cellCountperCluster <- data.frame(id = Idents(combined))
barplot = ggplot(data=cellCountperCluster, aes(x=id)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  theme_minimal()
barplot + labs(x="Cluster", y = "Cell count")
```

Cluster 18 & 14 are likely to be HC cluster

```{r}
cells.use <- WhichCells(combined, idents = c('16', '11'))
DimPlot(combined, reduction = "umap", group.by = "ident", cells.highlight = cells.use, sizes.highlight = 0.3) + NoLegend()
```
